# Workshop 1 Navigating R

2023 October 13

```{r}
#Do some stuff
#Without "#" symbol it will produces error because "Do some" is not a function or an assigned object

5*5 #Simple calculator stuff
```

## Storing Data Objects

```{r}
x <- 24 # <- symbol assigns integer value of 24 to x
y <- 3 # assigns 3 to y

x*y # equals 72

z <- x*y
z # running this will show the value assigned to z and also =72
```

## Activity 1 - Calculating mph, mpg, etc

How many miles per hour did I drive on a road trip?

What was the mpg?

```{r}
miles <- 1000
hours <- 15
mph <- miles/hours
mph

#Imaginary miles per gallon for car, maybe a prius
mpg_prius <- 80
mpg_truck <- 15

#kilometers per hour conversion
kmph <- mph*1.6
kmph

# In an imaginary view of the past
AvgGasCost <- 3.00

# gallons of gas consumed
gal_prius <- miles/mpg_prius
gal_truck <- miles/mpg_truck

# cost of gas per gallon
TripCost_prius <- gal_prius*AvgGasCost
TripCost_truck <- gal_truck*AvgGasCost
TripCost_prius 
TripCost_truck

```

## Encountering Errors

Many will occur overtime, but

-   object 'galons' not found

    -   missing an "l" in gallons

-   Commas often misplaces, or other symbols such as parenthesis etc.

## Vectors

-   a one dimensional object

-   Data Types

    -   integer

        -   whole numbers

    -   numeric

        -   numbers with decimals

    -   logical

        -   boolean TRUE or FALSE

    -   character

        -   words

```{r}
#what if we were tracking grades in a class?
# c() is the function concatenate and combines vectors into a group, basically like a column of data in a spreadsheet

grades <- c(94, 89, 100, 92) # Random grades for four students

# How to find the first value in the vector?
# use square brackets for this []
grades[1] # pulls the first element of a vector
grades[4] # pulls the fourth element of a vector
grades[1:3] # pulls the first through the third vecors

# We want to know the average test score
# we use the mean() function and type grades between the parenthesis
mean(grades)

# What if we want to figure out what a function does?
# We will put a ? before the name of the function
# This will open up the help area in the bottom right box in RStudio
?mean

# What if we want to know the minimum and maximum values?
min(grades)
max(grades)
```
